---
title: "Draft practical: "
subtitle: "subtitle"
author: "Wouter van Amsterdam"
eval: false
bibliography: references.bib
format: 
  html:
    toc: true
    self-contained: true
    code-fold: true
    df-print: paged
    callout-appearance: simple
    callout-icon: false
---

In this practical, ...


::: {.callout-tip appearance="default" icon=true}
The answers to each exercise are available as a collapsed `code` block. Try to work out the answer yourself before looking at this code block!
:::

First we load a package

```{r}
#| label: setup
#| message: false
#| warning: false
#| code-fold: false
#| eval: true

library(survival)
```

In this practical, we will also use the following two packages:

::: {.callout-note}
## Exercise 1
Grab a coffee
:::


::: {.callout-note}
## Exercise 2
Load the `mtcars` dataset in R 

```{r}
#| label: data-true
#| eval: true

data(mtcars)
```

:::

## Non-coding questions

::: {.callout-tip collapse="true"}
## Question Foo

answer: bar

:::

<!--Here's a cool fontawesome icon {{< fa thumbs-up >}}-->

::: {.callout-goals}
## my goals
:::


## Conclusion

You learned ABC

## Further reading

Read Chapter 9 of Pearls Causality [@pearlCausalDiagramsIdentification2009]

