---
title: Structural Causal Models
subtitle: introduction
author: Wouter van Amsterdam
date: 2024-08-06
format: 
    revealjs:
        theme: umcu.scss
        incremental: true
        width: 1600
        height: 900
        logo: umcu_blue.png
categories:
    - SCM
    - day2
bibliography: bibliography.bib
---

## Causal inference frameworks

### What are they for?

#### Mathematical language to
  - define *causal* quantities
  - express *assumptions*
  - derive how to *estimate* causal effects

## Causal inference frameworks

### Why learn more than one?

<!--todo: add hyperlink to day 1 materials-->

- On day 1 we learned about the Potential Outcomes framework
    - Defines causal effects in terms of (averages of) *individual potential outcomes*
    - Estimation requires assumptions of (conditional) exchangeability and positivity / overlap
- There isn't only 1 way to think about causality, find one that '*clicks*'
- Now we will learn another framework: *structural causal models* (SCMs)
    - SCM approach is broader in that it can define more different types of causal questions
- Equivalence: given the same data and assumptions, get the same estimates

## topics

- SCMs as computer programs
  - intervention as change in program
    - perfect versus soft intervention
- DAGs
  - variable types: confounders, mediators, colliders
- d-separation
- identifyability
  - back-door criterion
  - follow rules of do calculus
- bonus queries:
  - counterfactuals
  - probability of necessity, probability of sufficiency

## The world as a computer program

### Think of the world as a computer program with a set of 

- (endogenous) *variables*:
  - `surgery` = duration of surgery (hours)
  - `los` = length of stay in hospital post surgery (days)
  - `survival` = survival time (years)
- *background variables* (exogenous):
  - `u_surgery`
  - `u_los`
  - `u_survival`
- *functions* `f_` for each *variable* which depend on its *parents* `pa_` and its own *background* `u_`:
  - `surgery = f_surgery(pa_surgery,u_surgery)`
  - `los = f_los(pa_los, u_los)`
  - `survival = f_survival(pa_survival, u_survival)`
  
:::{.fragment}

<!-- The variables $\mathbb{V}$, background $\mathbb{U}$ and functions $\mathbb{F}$  -->
Together these define a *Structural Causal Model* [see definition 7.1.1 in @pearlLogicStructureBasedCounterfactuals2009]

:::

## Structural Causal Model 1

```{r}
#| label: scm1
#| execute: false
#| echo: true
#| code-line-numbers: "1-3|5-7|9-11|13-18|20"
#| output-location: fragment

f_surgery <- function(u_surgery) { # pa_surgery = {}
  u_surgery
}

f_los <- function(surgery, u_los) { # pa_los = {surgery}
  surgery + u_los
}

f_survival <- function(surgery, los, u_survival) { # pa_survival = {sugery, los}
  survival = los - 2 * surgery + u_survival
}

model1 <- function(u_surgery, u_los, u_survival) {
  surgery = f_surgery(u_surgery)
  los = f_los(surgery, u_los)
  survival = f_survival(surgery, los, u_survival)
  c(surgery=surgery, los=los, survival=survival)
}

model1(2, 1, 5)
```

## Submodel and Effect of Action

Take `model1`, replace `f_los` with a specific *value* (e.g. `7` days):

. . .

```{r}
#| label: submodel1
#| execute: false
#| echo: true
#| code-line-numbers: "1|3|8"
#| output-location: fragment

submodel7 <- function(u_surgery, u_los, u_survival) {
  surgery = f_surgery(u_surgery)
  los = 7
  survival = f_survival(surgery, los, u_survival)
  c(surgery=surgery, los=los, survival=survival)
}

submodel7(2, 1, 5)
```

. . .

This results in a *submodel* (`submodel7`) which is the *effect of action*

$$
 \text{do}(\text{los}=7)
$$

Denoted with the so-called *do-operator*

. . .

Intuitively this can be interpreted as *leave everything the same in the world,
but set this variable `los` to this value `7`*

## recursive Structural Causal Models have a unique implied Directed Acyclic Graphs

- note that `model1` is *recursive*, i.e. *acyclic*, meaning when following the chain of parents, you will never end up at the same variable twice
- each acyclic SCM implies a unique Directed Acyclic Graph
- for the remainder, we assume recursiveness



# SCM vs potential outcomes

- definition of causal effect (PO: unobserved counterfactuals, SCM: DAG)

## References

